Learning Augmented Joint-Space Task-Oriented Dynamical Systems: A Linear Parameter Varying and Synergetic Control Approach

### 2018-05-15 paper/reply_reviews/Reply_to_reviews.tex

\documentclass{article}
\usepackage[left=2cm, right=2cm, top=2cm]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{color}
\usepackage{graphicx}
\usepackage{subcaption}
\title{Response to Reviewers: Manuscript ID 17-1142.1: ``Learning Augmented Joint-Space Task-Oriented Dynamical Systems: A Linear Parameter Varying and Synergetic Control Approach"}

%\author{Your Name  \\
%	Your Company / University  \\
%	\and 
%	The Other Dude \\
%	His Company / University \\
%	}
%
%\date{\today}
% Hint: \title{what ever}, \author{who care} and \date{when ever} could stand 
% before or after the \begin{document} command 
% BUT the \maketitle command MUST come AFTER the \begin{document} command! 
\begin{document}
\maketitle
\small
First of all, we would like to thank all reviewers for the time they spent reviewing our paper. We greatly appreciate the comments, suggestions and opinions that we received. We hope to have addressed all their comments and suggestions satisfactorily.\\

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\noindent \underline{Comments from Reviewer \#1 (Reviewer: 4 (Review30069)):}
\begin{enumerate}
\item \textcolor{blue}{\textit{``As mentioned by the authors, there are many reasons to plan in joint space rather than task space. One of the important reasons to me seems to be able to plan end-effector orientation, not just position. Traditional movement primitives, for example DMPs, do not explicitly encode planning in the SO(3) space of orientations. There is research that addresses this question, for example [1], but they lose the simplistic appeal of movement primitives. I think it would be a great addition to this paper if the authors illustrate if their joint space DS can be used to accomplish orientation sensitive tasks, and generalize to end-effector orientations different that what was demonstrated.[1] Ude, Ale, et al. "Orientation in cartesian space dynamic movement primitives." Robotics and Automation (ICRA), 2014 IEEE International Conference on.}}"\\
\texttt{Response:}  We highly appreciate this comment. When we initially proposed this approach, we did not think about tackling the problem of reaching a target in orientation as well. Since our DS is learned in joint-space, indeed, we do not have to deal with the challenges that arise when trying to reach a target in both position and orientation as we do not have the constrains imposed by SO(3). There was theoretically no modification needed in our controller:
\begin{equation}
\underbrace{\dot{q}}_{\mathbb{R}^{m\times 1}} = -\underbrace{\mathcal{A}(q)}_{\mathbb{R}^{m\times m}}\underbrace{J^{T}(q)}_{\mathbb{R}^{m\times d}}\underbrace{(H(q)-x^*)}_{\mathbb{R}^{d\times 1}}
\label{eq:jtds}
\end{equation}
Originally, when working solely in position $d=3$, now we represent the target with $d=9$, where the first 3 dimensions correspond to Cartesian position and the remaining 6 correspond to the first and second columns of a rotation matrix $R \in SO(3)$. This only modifies the Jacobian $J(q)$ which is now $J(q)\in \mathbb{R}^{9 \times 7}$ rather than $J(q)\in \mathbb{R}^{3 \times 7}$. The role of $\mathcal{A}(q)$ remains  the same. The following paragraph has been incorporated, as well as some additions in the abstract and problem statement:
\textcolor{blue}{\small ``Finally, by learning a behavior in joint-space we can inherently reach a task-space target, not only in the Cartesian 3-D space $\mathbb{R}^3$, but also in the space of orientations $SO(3)$. Planning or learning the rotational component of motion is a challenging problem. Representing an orientation as a vector in Euclidean space may lead to inaccurate and unstable motions, due to its directional nature and vulnerability to singularities. Several works have proposed tailor-made learning approaches that consider the non-Euclidean geometry of the $SO(3)$ space to generate rotational motion [18][19][20]. These approaches, however, require an explicit coupling between position and orientation, that might cause discontinuities in the resulting end-effector motion. Such coupling is not necessary if the motion is encoded in joint-space."} With the included references being:\\
$[18]$ S. Kim, R. Haschke, and H. Ritter, “Gaussian mixture model for 3-dof orientations,” \textit{Robotics and Autonomous Systems}, vol. 87, pp. 28-37, 2017.\\
$[19]$ A. Ude, B. Nemec, T. Petri, and J. Morimoto, “Orientation in cartesian space dynamic movement primitives,” in \textit{2014 IEEE International Conference on Robotics and Automation (ICRA)}, May 2014, pp. 2997-3004.\\
$[20]$ M. J. A. Zeestraten, I. Havoutis, J. Silvrio, S. Calinon, and D. G. Caldwell, “An approach for imitation learning on riemannian manifolds,” \textit{IEEE Robotics and Automation Letters}, vol. 2, no. 3, pp. 1240-1247, July 2017.

Furthermore, we have re-learnt all demonstrated behaviors incorporating orientation in the task-space target $x^*$. Table I has been updated accordingly, showing no deterioration in generalization and performance. Finally, as will be discussed in the response to the Comments 10 and 12, we compare our approach to a Task-Space DS learnt in both position+orientation, showing a considerable improvement in accuracy while reaching the target.

\item \textcolor{blue}{\textit{``In the introduction, the authors stress that jacobian psuedo inverse
can be computationally expensive. Is this really true for a 7-dof robot
arm? Can the authors provide some numbers where this would really be an
issue for a robot controlled at 1000Hz? I find the argument that
jacobian inverse does not work for singularities and needs a set of
heuristics to be more compelling. Maybe the authors should rewrite the
introduction to stress more on this issue?"}}\\
\texttt{Response:} \small For a 7-DOF robot it is indeed negligible. Computational burden only becomes a real issue for higher degrees of freedom, like a multi-fingered hand or multi-arm robotic setting. We removed this remark in order to allow space to discuss about other issues. \\

\item \textcolor{blue}{\textit{``In Section II, criterion I, the authors talk about a positive symmetric matrix P. I do not see it in the
equations above. Is this a weighing matrix?"}}\\
\texttt{Response:} \small Removed, that was a carry-over from a previous formulation.\\

\item \textcolor{blue}{\textit{``In criterion II, the error is defined with respect to the velocity error, rather than a position or a combination
of both. This means that the motion can be different from the demonstrated motion. I understand
that this is important to keep the later optimization convex, but is there any disadvantage to using
velocity error rather than position error?"}}\\
\texttt{Response:} \small Optimizing for position distance alone would have the effect of learning a motion that converged to the average of previous motions, thus not changing the motion in new situations or fitting the joint velocity profile that is often a desired component of the motion. Using velocity error does mean that, for motions strongly dependent on joint position rather than joint velocity, the generated motions may generalize from the demonstrations in unexpected ways. However, as you mentioned, minimizing position error would, depending on the formulation, be prohibitively difficult. This would imply simulating the learned motion forward in time to recompute the robot's positions, making the optimization much more complicated; this way, we can treat each point in the motion independently.\\

\item \textcolor{blue}{\textit{`I do not completely understand the significance of A(q). From Figure 3, it seems like A(q) helps resolve
redundancy, while achieving the task space goal. Do the authors have a mathematical (in constrast to
intuitive) explanation for this? Does A(q) somehow relate to the null-space of the jacobian?"}}\\
\texttt{Response:} \small A(q) transforms a joint velocity vector approximately aligned with the task-space error into a different joint velocity vector that remains approximately aligned with the error, but which we want to choose to more-closely approximate our demonstrations. In this sense, it is resolving redundancy between all the possible joint velocity vectors that would move the controller closer to the origin. The paragraph in Section III has been modified to improve clarity. A(q) doesn't relate to the null-space of the Jacobian in any clear way, as its value can and does alter the task-space trajectory as well as the joint space trajectory. The only property that A(q) preserves is the joint velocity's moving the task space position closer to the goal, but the direction of motion is not necessarily the same as the direction of task-space error.\\

\item \textcolor{blue}{\textit{``In Section III, ?One can intuitively understand the control law as follows:$(H(q) - x^*)$ denotes the
position error w.r.t. the task target and by multiplying that error by the transposed Jacobian JT (q),
the error is projected into joint space (similar to Jacobian transpose control)?. I find this explanation
misleading as jacobian multiplied to position error does not project the error into joint space. Jacobian
transforms velocities, not positions. Maybe the law can be seen more as a first order approximation
of the error in task space. In any case, I would either clarify this statement more or take it out of the
manuscript all together. It seems like the DS is just a PD law that provably converges to a fixed attractor
in task space."}}\\
\texttt{Response:} \small The passage has been rewritten to remove the word "project", which you appropriately noted was incorrect, and further clarifies the nature of the DS as a PD controller. We have added the following text in the manuscript: 
\textcolor{blue}{\textit{``One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t. the task target, which we then reinterpret as the task-space velocity of a proportional controller. By multiplying that error by the transposed Jacobian $J^T(q)$, it is transformed into a joint-space velocity vector correlated with the error (similar to Jacobian transpose control [23][27]), see Fig. 2. The positive definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig. 3 illustrates the effects of  $\mathcal{A}(q)$ on the generated motion. Thus the controller can be thought of as a proportional controller in joint space.}}\\

\item \textcolor{blue}{\textit{``About figure 2 again, the position error in task and joint space are not simply multiplied by a Jacobian.
Although, task space error can be first-order approximated using a Jacobian multiplication with
joint space error. What does the right figure in Fig 2 exactly represent?"}}\\
\texttt{Response:} \small The figure on the left is simply the cartesian error in task space. The figure on the left is the Lyapunov potential function on which the jacobian transpose controller descends. In showing the attractive regions in joint-space corresponding to this Lyapunov function, we aim to offer insight into how the controller reformulates the task-space potential into a different yet related joint-space potential. The caption has been rewritten to add clarity.\\

\item \textcolor{blue}{\textit{``In Section IV, the authors say that they use EM to determine the parameters of the GMM in Eq 6. Can
they elaborate a little on this? What is the expectation taken over and what likelihood is maximized?"}}\\
\texttt{Response:} \small The algorithm computes the expected probability of generating each of the joint positions given the current values of the GMM's means, variance matrices, and scaling values. The algorithm then chooses GMM parameters to maximize the posterior likelihood of the model, using previously-computed expected component membership as likelihood weights when computing the new mixture. Specifically, we used MATLAB's \textit{fitgmdist} implementation. Added a citation to clarify.\\

\item \textcolor{blue}{\textit{``The learned dimensionality reduction and parameters seem task dependant. I wonder if the authors
have any idea on how to generalize this across tasks."}}\\
\texttt{Response:} \small The dimensionality reduction could be generalized across tasks that had the same joint postural synergies (i.e. same lower dimensional space) - that is, for example, all tasks in which the shoulder and elbow always move together. Similarly, one could imagine several similar motions that could be decomposed into shared sub-motions, thus allowing the parameters to be shared. However, in general, fixing some of the parameters while optimizing will simply lead to a less-flexible model, and considering that there is no intuitive scheme for \textit{a-priori} identifying motions with strongly similar sub-motions, we did not explore this question further. Given that the local behavior regions are defined by the GMM components, one could use the likelihood (or KL divergence) to determine the similarity of the motions when face with a new task. To answer your question directly, if one wanted to, one could reuse the learned dimensionality reduction for a new task which you believed to be similar, or copy certain components of a previous motion's GMM into the new GMM and leave them fixed throughout the optimization.

\item \textcolor{blue}{\textit{``The experiments have the same task-space goal as the
demonstrations. Can the authors include experiments that have goals
different from demonstrated, especially in orientation?"}}\\
\texttt{Response:} \small Yes. As we have shown in Appendix I, regardless of the target location (position/orientation), the motion generated by the proposed dynamical system asymptotically converges to the desired target. However, it is important to note that as (8) in Appendix I is semi-negative-definite and not negative-definite, there might be joint configuration $ (q )$ which result in $ J(q)\sum\limits_{k=1}^{K}\theta_k(q)A_kJ^T(q)=0 $. In this situation, the robot is stuck in a deadlock and, consequently, cannot reach the desired target.

\item \textcolor{blue}{\textit{``In similar light, maybe authors could present online adaptation of
A(q) to generalize to situations different than demonstrated, for
example a slightly moved obstacle?"}}\\
\texttt{Response:} \small The reviewer has pointed to a very interesting topic and the authors appreciate the reviewer’s suggestion. In this paper, we assumed that the learned model (namely $ \pi $'s, $ \mu $'s, $ \Sigma $'s and $A$'s) does not need to be updated or modified during the task execution. Hence, this would be out of the scope of this paper and we are planning to address it in future works.

\item \textcolor{blue}{\textit{``In table II, instead of the tracking error, maybe the more useful
error to see would be the error with respect to the plan or
demonstartion. One would assume that the proposed approach would be
worse at it than SEDS+IK because the minimized error is with respect to
the velocity and goal, not the position of the movement."}}\\
\texttt{Response:} \small This is, in general, not true. Based on the robot's motion and its configuration, there might be cases where the motion generated by SEDS does not even reach the target. To clarify this, we have systematically studied the performance of the motion generated by JT-DS and SEDS+IK in terms of (i) reaching to the desired target and (ii) following the demonstrated end-effector motions, See Figure \ref{fig:SEDSvsJTDS}. We used the data-sets from the seven tasks mentioned in Section V.A to construct SEDS and JT-DS. Both methods are tested by the same the target and the initial configurations, which were extracted from the demonstrations. We define $ x^*=[x^*_{ori};x^*_{pos}] $ as the target configuration that includes the desired position as well as the desired  orientation of the end-effector. The motion generated by SEDS is converted into the 7-DOF joints state using the damped least square IK solver.\\ 
In Figure \ref{fig:SEDSvsJTDSP:E} and Figure \ref{fig:SEDSvsJTDSO:E}, the error between the final location of the robot and the target is illustrated. As it can be seen, the motion generated by the proposed dynamical system reaches the target, whereas, the motion generated by SEDS does not; regardless of the fact that both models are theoretically stable. The main cause of failure for SEDS is due to the IK solver. As SEDS generates the motion of the robot without considering its joint configurations, the robot might reach its joint limits.  Whereas, as JT-Ds directly shapes the joints' motion, by learning the desired joint-space behavior, one can avoid these situations.

In Figure \ref{fig:SEDSvsJTDSP} and Figure \ref{fig:SEDSvsJTDSO}, the error between the demonstrated and the excused motions are illustrated. As it can be seen, even though JT-DS is not tailored for following the end-effector motions, its performance is comparably better than SEDS.

Although including this study would highlight one of the advantages of using JT-DS over Cartesian-motion-generators, as this does not highlight the main focus of this paper and due to the space limitations, we have decided to not include this in the paper.

\begin{figure}[!t]
 \hspace*{-2.4cm}
	\begin{subfigure}[t]{0.63\linewidth}
	\centering
		\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_P_End.pdf}
		\caption{}
		\label{fig:SEDSvsJTDSP:E}
	\end{subfigure}%
	\hspace{-0.2cm}
	\begin{subfigure}[t]{0.63\linewidth}
	\centering
		\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_O_End.pdf}
		\caption{ }
		\label{fig:SEDSvsJTDSO:E}
	\end{subfigure}\\		 
	\hspace*{-2.4cm}\begin{subfigure}[t]{0.63\linewidth}
			\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_P.pdf}
		\caption{ }
			\label{fig:SEDSvsJTDSP}
		\end{subfigure}%
		\begin{subfigure}[t]{0.63\linewidth}
			\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_O.pdf}
			\caption{ }	
			\label{fig:SEDSvsJTDSO}
		\end{subfigure}\\
	\caption{The performance comparison between SEDS+IK and JT-DS while the initial and the target points are same. The error between the orientations is the difference between the first two columns of the corresponding rotation matrix representation.}
	\label{fig:SEDSvsJTDS}
\end{figure}

\end{enumerate}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\newpage 
\underline{Comments from Reviewer \#2 (Reviewer: 5 (Review30371)):}
\begin{enumerate}
\item \textcolor{blue}{\textit{``My first concern regards you choice of representing the demonstrated
movement as a sequence of postures and hence looking for synergies by
using PCA of K-PCA analysis. Why not using functional PCA (f-PCA)
analysis and determine the few moving synergies that make possible the
execution of the demonstrated behaviours? It would be that, when you
combine the local synergies in space and time by using functions
$\theta_k(q)$, you get what you might obtain by using directly f-PCA.
However, the last one are valid for the whole movement and can be
combined by constant coefficients."}}\\
\texttt{Response:}  See below.

\item \textcolor{blue}{\textit{``I also found the paper lacking of physical intuitions, especially in	the experimental part, about the entities introduced in previous sections. In particular, what is the lesson we learn after your analysis of synergies and the obtained lower dimensional space for the few movements shown in section V? Is it possible to extract some global	synergic behavior? Moreover, in section IV.A authors introduce the function phi(q) named the embedding that maps the joint configuration into a lower dimensional space. Is it associated to a whole behaviour or, as matrices $A_k$, has a local validity? I lost this point. Can you show/report or describe what it represents in the joint space for one or more cases in Section V? What is this lower dimensional space for the cases in Section V?"}}\\
\texttt{Response:}  See below.

### 2018-05-15 paper/final-source-for-zip/main_RA-L_final.tex
Final update, folder restructure, final proof of published paper

%\documentclass[9pt,technote,fleqn]{IEEEtran}
\documentclass[letterpaper, 10 pt, conference,fleqn]{ieeeconf}

@@

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
%\title{Learning Joint-Space Stable Dynamical Systems for Task-Space Objectives}
\title{Learning Augmented Joint-Space Task-Oriented Dynamical Systems: \\ A
Linear Parameter Varying and Synergetic Control Approach}

@@

%
%
\author{Yonadav~Shavit*, Nadia~Figueroa*, Seyed~Sina~Mirrazavi~Salehian*, Aude~Billard
% <-this % stops a space
\thanks{*These authors contributed equally to this work.}
\thanks{Y. Shavit is with the Department
of Electrical Engineering and Computer Science, Massachusetts Institute of Technology, Cambridge,
MA, 02139 USA. E-mail: yonadav.shavit@gmail.com.}% <-this % stops a space
\thanks{N. Figueroa, S. S. M. Salehian, and A. Billard are with the Swiss Federal Institute of Technology (EPFL), 1015 Lausanne, Switzerland. \\E-mail: \{nadia.figueroafernandez,sina.mirrazavi,aude.billard\}@epfl.ch.}
}% <-this %

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

@@

%\begin{IEEEkeywords}
%Dynamical Systems, Imitation learning, Robot Kinematics, Gaussian Mixture Models
%\end{IEEEkeywords}

\IEEEpeerreviewmaketitle
\vspace{-10pt}
\section{Introduction}
\label{sec:intro}
Robot motion planning in joint-space has long been a major field of study \cite{kelly2006control}. For manipulation problems with an objective defined in task-space, we can often find a myriad of joint-space trajectories to achieve the same task-space goal. In many cases, however, certain joint-space trajectories are favored over others; for example, when we expect the robot to follow a desired joint-space behavior or ``style", as illustrated in Fig. \ref{fig:robot_example}. In this paper, we will explore the problem of learning a preferred joint-space behavior from previously demonstrated trajectories while still accomplishing a task-space goal, through the Learning from Demonstrations (LfD) paradigm \cite{billard2008robot,Argall:RAS:2009}. Most LfD approaches learn motions either in joint space
\cite{Garrido:Neuro:2015,Schaal:AI:2002} or in task space
\cite{Pastor:ICRA:2009,Calinon:ISR:2015} via probabilistic models.
To ensure that the learnt models reach the desired task-space target, Dynamical System (DS) formulations have been used to ensure stability and  convergence \cite{khansari2011learning, Neumann:RAS:2015}.

In many cases, however, one may want to specify a joint-space behavior that the physical robot should follow, while simultaneously reaching a task-space objective. For example, learned joint-space behaviors are helpful in playing ping-pong \cite{huang2016jointly}, grasping \cite{calinon2010learning}, and avoiding self-collisions of bi-manual manipulators \cite{silverio2017learning}. For a visual example of a task-space problem in which joint behavior is essential, see Fig.\ref{fig:robot_example}. Moreover, learning a motion in joint space allows us to avoid inverse kinematic (IK) approximations. Task-space motion generators, as the ones learned from demonstrations, all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse IK approximations and variants thereof \cite{kelly2006control}. When the main focus is on executing a specific task-space behavior, regardless of a joint-space constraint, this approach is sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}. However, for other applications, such an approach yields significant problems \cite{buss2004introduction}. Mainly, when the Jacobian matrix cannot be inverted (i.e. when the robot is near a singularity) its behavior becomes erratic, requiring layers of additional engineering to generate smooth trajectories and ensure the desired task-space behavior. This encapsulates the main source of inaccuracies in tasks that require fast dynamical motions, such as catching/reaching for moving objects \cite{7439839,Salehian-RSS-16}. 
%Defining a controller in joint-space would avoid IK and all its associated drawbacks.

Finally, by learning a behavior in joint-space we can inherently reach a task-space target, not only in the Cartesian 3-D space $\mathbb{R}^3$, but also in the space of orientations $SO(3)$. Planning or learning the rotational component of motion is a challenging problem. Representing an orientation as a vector in Euclidean space may lead to inaccurate and unstable motions, due to its directional nature and vulnerability to singularities. Several works have proposed tailor-made learning approaches that consider the non-Euclidean geometry of the $SO(3)$ space to generate rotational motion \cite{KIM201728} \cite{6907291} \cite{7829369}. These approaches, however, require an explicit coupling between position and orientation, that might cause discontinuities in the resulting motion. Such coupling is not necessary if the motion is encoded in joint-space. 
\begin{figure}[t]
\centering
\includegraphics[scale=.16,trim={0 0 0 0cm},clip]{../Without_Obstcle_white_new.png}\includegraphics[scale=.135,trim={0 0 0 0cm},clip]{../With_Obstcle.png}
\caption{Two robot motions in \textcolor{orange}{\textbf{joint-space}} accomplishing \textit{similar} behavior in \textcolor{cyan}{\textbf{task-space}} (pouring chips in a bowl). The left example avoids a known obstacle in its workspace, while the right one does not.}
\label{fig:robot_example}
\vspace{-20pt}
\end{figure}

Several approaches have tackled the problem of learning joint-space behaviors with task-space objectives. \cite{calinon2008probabilistic} notably attempted to address this problem by learning separate motion policies in task space and the null space of the Jacobian (which would not affect task-space position), and driving the robot with a weighted sum of the two.  However, this approach does not seek convergence to the desired task-space target and is still reliant on computing the pseudo-inverse Jacobian. \cite{calinon2010learning} and \cite{silverio2017learning} expand on this approach by projecting task-space constraints into joint space using IK, and then learning a joint-space policy that incorporates both task and joint space constraints, but this similarly relies on IK approximations and does not ensure convergence to an attractor.
% FACT-CHECK: is this what Calinon 2009 and Silverio et al. 2017 are doing? They don't ensure convergence, right? TP-GMM never seems to address it ->CORRECT
%\cite{lee2014unifying} learns from demonstration in both joint and task space within the ``thin-plate spline" trajectory warping framework. % but does not propose a dynamical system for generating motions. 
\cite{hersch2008reaching} proposed an approach with similar properties to our desiderata, where two concurrent DS, one in task-space and one in joint-space, are modulated by enforcing kinematic coherence constraints to avoid singularities. The resulting DS avoids singularities through generalization of the pseudo-inverse approximations. However, because the two DS have their own unique attractors, the non-linear interaction between them imposed by the kinematic constraints does not ensure that the combined DS has a unique attractor. This gives rise to spurious attractors or cycles, and thus must be carefully tuned in order to avoid them.

In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations. To approximate this DS, we further propose an algorithm to learn a set of \emph{behavior synergies}, each of which corresponds to a different stable behavior in joint space, and modulates the use of these synergies throughout joint-space using a learned Linear Parameter Varying (LPV) system. We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions. We then learn a policy in embedding space for combining our behavior synergies to accurately reconstruct the demonstrated trajectories. 
Hence, our dynamical system:
\begin{enumerate}[leftmargin=*]
\item computes a \begin{bf}motion in joint-space\end{bf} that provably and asymptotically converges to a \begin{bf}task-space \end{bf}target.
\item is formulated such that \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations in \textbf{synergy space}.
\item can transit through kinematic \textbf{singularities}.
\end{enumerate}

The most similar approach to our proposed DS is the Jacobian transpose (JT) control method \cite{wolovich1984computational} . The JT control is an IK method that yields a dynamical system in joint space
which converges stably over time to a desired end-effector target, without the need for pseudo-inverse computations. It shares some of our approach's advantages: fast computation and provable task-space stability. However, despite some previous work designing velocity adjustments by hand \cite{Shi2016}, this is, to the best of our knowledge the first work to employ a JT system to learn behaviors from demonstrations. Furthermore, by formulating our LPV system on a latent embedding (via dimensionality reduction schemes), we are capable of discovering meaningful local behavior synergy regions, while being robust to outliers, noise and redundancies that might arise from raw demonstrations. This leads to a compelling improvement in generalization of the demonstrated behavior, as opposed to learning the LPV system solely in joint space.

>>>>

%\documentclass[9pt,technote,fleqn]{IEEEtran}
\documentclass[letterpaper, 10 pt, journal, twoside, fleqn]{IEEEtran}

@@

\begin{document}

% Paper headers 
\markboth{IEEE Robotics and Automation Letters. Preprint Version. Accepted April, 2018}
{Shavit \MakeLowercase{\textit{et al.}}: Learning Augmented Joint-Space Task-Oriented Dynamical Systems}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
%\title{Learning Joint-Space Stable Dynamical Systems for Task-Space Objectives}
\title{Learning Augmented Joint-Space Task-Oriented Dynamical Systems: A
Linear Parameter Varying and Synergetic Control Approach}

@@

%
%
\author{Yonadav~Shavit$^{*1}$, Nadia~Figueroa$^{*2}$, Seyed~Sina~Mirrazavi~Salehian$^{*2}$, Aude~Billard$^{2}$
% <-this % stops a space
\thanks{Manuscript received: November 21, 2017; Revised March 16, 2018; Accepted April 21, 2018.}%Use only for final RAL version
\thanks{This paper was recommended for publication by Editor Han Ding upon evaluation of the Associate Editor and Reviewers' comments. 
This work was supported by EU project Cogimon H2020\textendash ICT\textendash 23\textendash2014.}
\thanks{$^*$These authors contributed equally to this work.}
\thanks{$^{1}$Y. Shavit is with the Department of Electrical Engineering and Computer Science, Massachusetts Institute of Technology, Cambridge,
MA, 02139 USA {\tt\footnotesize yonadav.shavit@gmail.com}}% <-this % stops a space
\thanks{$^{2}$ N. Figueroa, S. S. M. Salehian, and A. Billard are with the Swiss Federal Institute of Technology (EPFL), 1015 Lausanne, Switzerland
{\tt\footnotesize\{nadia.figueroafernandez,sina.mirrazavi,aude.billard\}\\ \tt\footnotesize @epfl.ch}}%
\thanks{Digital Object Identifier (DOI): see top of this page.}%
}
% <-this %

\maketitle
%\thispagestyle{empty}
%\pagestyle{empty}

@@

\begin{IEEEkeywords}
Motion Control, Learning from Demonstration, Kinematics, Gaussian Mixture Models, Joint-Space Control
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle
\vspace{-10pt}
\section{Introduction}
\label{sec:intro}
\IEEEPARstart{R}{obot} motion planning in joint-space has long been a major field of study \cite{kelly2006control}. For manipulation problems with an objective defined in task-space, we can often find a myriad of joint-space trajectories to achieve the same task-space goal. In many cases, however, certain joint-space trajectories are favored over others; for example, when we expect the robot to follow a desired joint-space behavior or ``style", as illustrated in Fig. \ref{fig:robot_example}. In this paper, we will explore the problem of learning a preferred joint-space behavior from previously demonstrated trajectories while still accomplishing a task-space goal, through the Learning from Demonstrations (LfD) paradigm \cite{billard2008robot,Argall:RAS:2009}. Most LfD approaches learn motions either in joint space
\cite{Garrido:Neuro:2015,Schaal:AI:2002} or in task space
\cite{Pastor:ICRA:2009,Calinon:ISR:2015} via probabilistic models.
To ensure that the learnt models reach the desired task-space target, Dynamical System (DS) formulations have been used to ensure stability and  convergence \cite{khansari2011learning, Neumann:RAS:2015}.

In many cases, however, one may want to specify a joint-space behavior that the physical robot should follow while simultaneously reaching a task-space objective. For example, learned joint-space behaviors are helpful in playing ping-pong \cite{huang2016jointly}, grasping \cite{calinon2010learning}, and avoiding self-collisions of bi-manual manipulators \cite{silverio2017learning}. For a visual example of a task-space problem in which joint behavior is essential, see Fig.\ref{fig:robot_example}. Moreover, learning a motion in joint space allows us to avoid inverse kinematic (IK) approximations. Task-space motion generators learned from demonstrations all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse IK approximations and variants thereof \cite{kelly2006control}. When the main focus is on executing a specific task-space behavior, regardless of a joint-space constraint, this approach is sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}. However, for other applications, such an approach yields significant problems \cite{buss2004introduction}. Mainly, when the Jacobian matrix cannot be inverted (i.e. when the robot is near a singularity) its behavior becomes erratic, requiring layers of additional engineering to generate smooth trajectories and ensure the desired task-space behavior. This encapsulates the main source of inaccuracies in tasks that require fast dynamical motions, such as catching/reaching for moving objects \cite{7439839,Salehian-RSS-16}. 
%Defining a controller in joint-space would avoid IK and all its associated drawbacks.

Finally, by learning a behavior in joint-space we can inherently reach a task-space target, not only in the Cartesian 3-D space $\mathbb{R}^3$, but also in the space of orientations $SO(3)$. Planning or learning the rotational component of motion is a challenging problem. Representing an orientation as a vector in Euclidean space may lead to inaccurate and unstable motions, due to its directional nature and vulnerability to singularities. Several works have proposed tailor-made learning approaches that consider the non-Euclidean geometry of the $SO(3)$ space to generate rotational motion \cite{KIM201728} \cite{6907291} \cite{7829369}. These approaches, however, require an explicit coupling between position and orientation, that might cause discontinuities in the resulting motion. Such coupling is not necessary if the motion is encoded in joint-space. 
\begin{figure}[t]
\centering
\includegraphics[scale=.155,trim={0 0 0 0cm},clip]{./figures/Without_Obstcle_white_new.png}\includegraphics[scale=.125,trim={0 0 0 0cm},clip]{./figures/With_Obstcle.png}
\caption{Two robot motions in \textcolor{orange}{\textbf{joint-space}} accomplishing \textit{similar} behavior in \textcolor{cyan}{\textbf{task-space}} (pouring chips in a bowl). The left example avoids a known obstacle in its workspace, while the right one does not.}
\label{fig:robot_example}
\vspace{-20pt}
\end{figure}

Several approaches have tackled the problem of learning joint-space behaviors with task-space objectives. \cite{calinon2008probabilistic} notably attempted to address this problem by learning separate motion policies in task space and the null space of the Jacobian (which would not affect task-space position), and driving the robot with a weighted sum of the two.  However, this approach does not seek convergence to the desired task-space target and is still reliant on computing the pseudo-inverse Jacobian. \cite{calinon2010learning} and \cite{silverio2017learning} expand on this approach by projecting task-space constraints into joint space using IK, and then learning a joint-space policy that incorporates both task and joint space constraints, but this similarly relies on IK approximations and does not ensure convergence to an attractor.
% FACT-CHECK: is this what Calinon 2009 and Silverio et al. 2017 are doing? They don't ensure convergence, right? TP-GMM never seems to address it ->CORRECT
%\cite{lee2014unifying} learns from demonstration in both joint and task space within the ``thin-plate spline" trajectory warping framework. % but does not propose a dynamical system for generating motions. 
\cite{hersch2008reaching} proposed an approach with similar properties to our desiderata, where two concurrent DS, one in task-space and one in joint-space, are modulated by enforcing kinematic coherence constraints to avoid singularities. The resulting DS avoids singularities through generalization of the pseudo-inverse approximations. However, because the two DS have their own unique attractors, the non-linear interaction between them imposed by the kinematic constraints does not ensure that the combined DS has a unique attractor. This gives rise to spurious attractors or cycles, and thus requires careful tuning to avoid them.

In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations. To approximate this DS, we further propose an algorithm to learn a set of \emph{behavior synergies}, each of which corresponds to a different stable behavior in joint space, and modulate the use of these synergies throughout joint-space using a learned Linear Parameter Varying (LPV) system. We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions. We then learn a policy in embedding space for combining our behavior synergies to accurately reconstruct the demonstrated trajectories.
Hence, our dynamical system:
\begin{enumerate}[leftmargin=*]
\item computes a \begin{bf}motion in joint-space\end{bf} that provably and asymptotically converges to a \begin{bf}task-space \end{bf}target.
\item is formulated such that \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations as \textbf{synergies}.
\item can transit through kinematic \textbf{singularities}.
\end{enumerate}

The most similar approach to our proposed DS is the Jacobian transpose (JT) control method \cite{wolovich1984computational} . The JT control is an IK method that yields a dynamical system in joint space
which converges stably over time to a desired end-effector target, without the need for pseudo-inverse computations. It shares some of our approach's advantages: fast computation and provable task-space stability. However, despite some previous work designing velocity adjustments by hand \cite{Shi2016}, this is, to the best of our knowledge the first work to employ a JT system to learn behaviors from demonstrations. Furthermore, by formulating our LPV system on a latent embedding (via dimensionality reduction schemes), we are able to discover meaningful local behavior synergy regions, while being robust to outliers, noise and redundancies that might arise from raw demonstrations. This leads to a compelling improvement in generalization of the demonstrated behavior, as opposed to learning the LPV system solely in joint space.
